---
title: "6450hw2"
author: "Ryan McKiernan"
date: "7/25/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lpSolve)
```

## Question 1

i) maximize after tax earnings, i.e. 

max(0.043a + (0.054)(0.5)b + (0.05)(0.5)c + (0.044)(0.5)d + 0.045e) s.t.

a + b + c + d + e <= 10,000,000

(2a + 2b + c + d + 5e) / (a + b + c + d + e) <= 1.4 

  = 0.6a + 0.6b - 0.4c - 0.4d + 3.6e <= 0

b + c + d >= 4,000,000

(9a + 15b + 4c + 3d + 2e) / (a + b + c + d + e) <= 5

  = 4a + 10b - c - 2d - 3e <= 0

```{r}
obj_fun1 = c(0.043,0.027,0.025,0.022,0.045)
constr_lhs = matrix(c(1,1,1,1,1,
                      0.6,0.6,-0.4,-0.4,3.6,
                      0,1,1,1,0,
                      4,10,-1,-2,-3),
                    ncol=5,byrow=TRUE)
constr_dir = c("<=","<=",">=","<=")
constr_rhs = c(10000000,0,4000000,0)
solution1 = lp("max",obj_fun1,constr_lhs,constr_dir,constr_rhs)
solution1
solution1$solution
```

Solution: The bonds that will be purchased are A, C, and E.


ii) Borrow up to 1 million at 5.5% before taxes. Here, we will add a
decision variable for the amount borrowed (z) that cannot exceed 1 million.
Besides that, we will add it to our initial monetary constraint which becomes

a + b + c + d + e - z <= 10,000,000

after taking into account taxes. With the rest of the constraints being the
same, our objective function becomes:

(0.043a + 0.027b + 0.025c + 0.022d + 0.045e - 0.0275z)

```{r}
obj_fun1b = c(0.043,0.027,0.025,0.022,0.045,-0.0275)
constr_lhs1b = matrix(c(1,1,1,1,1,-1,
                      0.6,0.6,-0.4,-0.4,3.6,0,
                      0,1,1,1,0,0,
                      4,10,-1,-2,-3,0,
                      0,0,0,0,0,1),
                    ncol=6,byrow=TRUE)
constr_dir1b = c("<=","<=",">=","<=","<=")
constr_rhs1b = c(10000000,0,4000000,0,1000000)
solution1b = lp("max",obj_fun1b,constr_lhs1b,constr_dir1b,constr_rhs1b)
solution1b
solution1b$solution
```

Solution: The same bonds are purchased, the full 1 million is used and there
is slightly more money made off of the investment.

iii) For the primal problem, our goal was to maximize the return on our
investment into the bonds. For the dual problem, were looking to minimize
the quality rating and years to maturity needed. By doing this, we
minimize the years to maturity and quality scale such that the gov/income
constraints are held.



## Question 2

Goal is to minimize cost of chemical production, which are rooted from
unit, storage, fixed, setup, and shutdown costs.

i) Decision variables: by month, we have:

(Jan/Feb/Mar/Apr/May/Jun) -> (a/b/c/d/e/f)

Unit costs: (4a+6b+8c+10d+12e+14f) determined by stock requirements,

Demand (units): (100a+150b+200c+150d+100e+300f)

Capacity (units): (250a+200b+150c+200d+150e+250f)

Storage costs depend on stock at end of month, as do the other costs.

Our objective function is:

min z = sum(c x q) + 2s + 700 (startup cost in Jan)
  
  z = 4a + 6b + 8c + 10d + 12e + 14f + 2u + 2v + 2w + 2x + 2y + 2z + 700
  
With demand and capacity constraints:

a-u=100, u+b-v=150, v+c-w=200, w+d-x=150, x+e-y=100, y+f-z=300,

a <= 250, b <= 200, c <= 150, d <= 200, e <= 150, f <= 250,

u,v,w,x,y,z >= 30

there is also a binary variable for the fixed costs where
if q >= 0 then j = 1 else j = 0


ii) With this regulation taking into effect, the problem is reformulated to
where for instance if April is shut down, we have:

min z = sum(c x q) + 2s + 700 + 1000

  z = 4a + 6b + 8c + 12e + 14f + 2u + 2v + 2w + 2x + 2y + 2z + 1700

where the demand and capacity constraints would be the same.

iii) With the extra fixed cost, we would have it where a binary variable 
would be added and would apply to the capacity constraints as well as add an
additional constraint that factors in the fixed cost.



## Question 3

west: 70K, midwest: 50K, east: 60K, south: 40K, invest at 20% return

50000/year to operate a lockbox

To figure out which lockboxes should be opened, we have to first figure out
the amount of loss thats associated with each assignment. This can be
obtained by taking the number of days x the daily value of each region
x the 20% investment rate. Doing that, we get the following numbers:

(LA/Chicago/New York/Atlanta):
West: (28K/84K/112K/112K)
Midwest: (60K/20K/50K/50K)
East: (96K/60K/24K/60K)
South: (64K/40K/40K/16K)

From there, we want to establish our objective function, which is to minimize
the cost of handling the checks. If we form our variables in a grid going
from a to p, we get:

min(28a+84b+112c+112d+60e+20f+50g+50h+96i+60j+24k+60l+64m+40n+40o+16p
    + 50w + 50x + 50y + 50z)

where w-z represent a lockbox being opened. With regards to constraints, we
have

a+b+c+d = 1

e+f+g+h = 1

i+j+k+l = 1

m+n+o+p = 1

a+e+i+m <= 100w

b+f+j+n <= 100x

c+g+k+o <= 100y

d+h+l+p <= 100z

sum(a:p) <= sum(w:z)

we also have constraints for each city-location combo where they are less than 
their respective to their respective lockboxes i.e.

a <= w , b <= x , e <= w and so forth. This gets us the following code:

```{r}
obj_fun3 = c(28,84,112,112,60,20,50,50,96,60,24,60,64,40,40,16,50,50,50,50)
constr_lhs3 = matrix(c(1,1,1,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,
                       0,0,0,0,1,1,1,1,0,0,0,0,0,0,0,0,0,0,0,0,
                       0,0,0,0,0,0,0,0,1,1,1,1,0,0,0,0,0,0,0,0,
                       0,0,0,0,0,0,0,0,0,0,0,0,1,1,1,1,0,0,0,0,
                       1,0,0,0,1,0,0,0,1,0,0,0,1,0,0,0,-1,0,0,0,
                       0,1,0,0,0,1,0,0,0,1,0,0,0,1,0,0,0,-1,0,0,
                       0,0,1,0,0,0,1,0,0,0,1,0,0,0,1,0,0,0,-1,0,
                       0,0,0,1,0,0,0,1,0,0,0,1,0,0,0,1,0,0,0,-1,
                       
                       1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,
                       0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,
                       0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,
                       0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,
                       0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,
                       0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,
                       0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,-1,0,
                       0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,-1,
                       0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,-1,0,0,0,
                       0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,-1,0,0,
                       0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,-1,0,
                       0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,-1,
                       0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,-1,0,0,0,
                       0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,-1,0,0,
                       0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,-1,0,
                       0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,-1
                       ),ncol=20,byrow=TRUE)
constr_dir3 = c("=","=","=","=","<=","<=","<=","<=",
                "<=","<=","<=","<=","<=","<=","<=","<=",
                "<=","<=","<=","<=","<=","<=","<=","<=")
constr_rhs3 = c(1,1,1,1,4,4,4,4,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)

solution3 = lp("min",obj_fun3,constr_lhs3,constr_dir3,constr_rhs3)
solution3
solution3$solution
```

Solution: (a,g,k,o) and the lockboxes that will be used are the first (LA) and
the second (New York)
