---
title: "MSBA 6451 HW1"
author: "Ryan McKiernan"
date: "7/19/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lpSolve)
```

## Question 1

Product 1=a, Product 2=b, Product 3=c, Product 4=d

Goal is to maximize profit while adhering to the various constraints for the 4 
products, which includes machine hours/assembly time, contract requirements, 
and government regulations.

Max(6a + 4b + 6c + 8d) s.t.

3a + 2b + 2c + 4d <= 480 (machine constraint),

a + b + 2c + 3d <= 400 (polish constraint),

2a + b + 2c + d <= 400 (assemble constraint),

a >= 50, b + c >= 100 (distributor constraints), d <= 25 (government constraint)

```{r}
obj_fun1 = c(6,4,6,8)
constr_lhs = matrix(c(3,2,2,4,1,1,2,3,2,1,2,1,1,0,0,0,0,1,1,0,0,0,0,1),
                    ncol=4,byrow=TRUE)
constr_dir = c("<=","<=","<=",">=",">=","<=")
constr_rhs = c(480,400,400,50,100,25)

solution1 = lp("max",obj_fun1,constr_lhs,constr_dir,constr_rhs)
solution1
solution1$solution
```

Solution: a=50 units, b=0 units, c=145 units, d=10 units, profit = $1,250


## Question 2

h=burgers, p=patties, m=meatloaf, r=round steak, c=chuck steak, k=pork

With this problem, we need to use up as many pounds of meat in order to minimize
the storage amount. This means maximizing the amount of product made in which
the objective function is created by adding up the percentages across the
products. From there the constraints are made up of the limits to how much meat
we have and how much is needed for each product, ignoring the inexpensive
nonmeat filler.

Max(0.7h + 0.7p + 0.8m) s.t.

0.2h + 0.1m <= 200 (round steak),

0.5h + 0.5p + 0.4m <= 800 (chuck steak),

0.2p + 0.3m <= 150 (pork),

h,p,m >= 0


```{r}
obj_fun2 = c(.7,.7,.8)
constr_lhs2 = matrix(c(.2,0,.1,.5,.5,.4,0,.2,.3,1,0,0,0,1,0,0,0,1)
                     ,ncol=3,byrow=TRUE)
constr_dir2 = c("<=","<=","<=",">=",">=",">=")
constr_rhs2 = c(200,800,150,0,0,0)

solution2 = lp("max",obj_fun2,constr_lhs2,constr_dir2,constr_rhs2)
solution2
solution2$solution
```

Solution: h = 937.5lbs, p = 562.5lbs, m = 125lbs where all the meat is used
and there is none left over.


## Question 3

r=regular, p=premium, d=domestic, f=foreign

For this problem, we need to break up the two categories into the 4 possible
combinations that are possible for selling gas. The objective function, which
is based in profit, can be created from (selling price – cost). The numbers for
vapor pressure and octane rating are obtained from 
(vapor pressure-max vapor pressure) and (octane rating – min octane rating).
The rest of the constraints are obtained from the numbers provided in the tables

Max(4rd -3rf + 6pd -pf) 	s.t. 

2rd – 8rf <= 0 (regular vapor pressure),

2pd – 8pf <= 0 (premium vapor pressure),

-rd + 10rf >= 0 (regular octane rating),

-6pd + 5pf >= 0 (premium octane rating),

Rd + rf <= 100,000 (max demand regular),

Pd + pf <= 20,000 (max demand premium),

Rd + rf >= 50,000 (min delivery regular),

Pd + pf >= 5,000 (min delivery premium),

Rd + pd <= 40,000 (domestic inventory),

Rf + pf <= 60,000 (foreign inventory),

Rd,rf,pd,pf >= 0

```{r}
obj_fun3 = c(4,-3,6,-1)
constr_lhs3 = matrix(c(2,-8,0,0,0,0,2,-8,-1,10,0,0
                      ,0,0,-6,5,1,1,0,0,0,0,1,1
                      ,1,1,0,0,0,0,1,1,1,0,1,0
                      ,0,1,0,1,1,0,0,0,0,1,0,0
                      ,0,0,1,0,0,0,0,1),ncol=4,byrow=TRUE)
constr_dir3 = c("<=","<=",">=",">=","<=","<=",">=",">=","<=",
                "<=",">=",">=",">=",">=")
constr_rhs3 = c(0,0,0,0,100000,20000,50000,5000,40000,60000,0,0,0,0)

solution3 = lp("max",obj_fun3,constr_lhs3,constr_dir3,constr_rhs3)
solution3
solution3$solution
```

Solution: rd=40,000 , rf=10,000 , pd=0 , pf=5,000 , weekly profit = $125,000
